---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

# censusaggregate

The goal of censusaggregate is to easily retrieve and aggregate Canadian census vectors, and their breakdowns, across multiple (shared) geographies.

## Installation

You can install censusaggregate from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("purposeanalytics/census-aggregator-package")
```

## API key and caching

censusaggregate uses the [`cancensus`](https://github.com/mountainMath/cancensus) package and [CensusMapper API](https://censusmapper.ca/). As per their documentation:

<blockquote>
You can obtain a free API key by [signing up](https://censusmapper.ca/users/sign_up) for a CensusMapper account. To check your API key, just go to "Edit Profile" (in the top-right of the CensusMapper menu bar). Once you have your key, you can store it in your system environment so it is automatically used in API calls. To do so just enter cancensus::set_api_key(\<your_api_key\>, install = TRUE).

CensusMapper API keys are free and public API quotas are generous; however, due to incremental costs of serving large quantities of data, there are some limits to API usage in place. For most use cases, these API limits should not be an issue. Production uses with large extracts of detailed geographies may run into API quota limits.

</blockquote>

In addition, censusaggregate uses `cancensus`' caching infrastructure to avoid repeat calls to the API:

<blockquote>

By default, cancensus caches in R's temporary directory, but this cache is not persistent across sessions. In order to speed up performance, reduce quota usage, and reduce the need for unnecessary network calls, we recommend assigning a persistent local cache using cancensus::set_cache_path(<local cache path>, install = TRUE), this enables more efficient loading and reuse of downloaded data. Users will be prompted with a suggestion to change their default cache location when making API calls if one has not been set yet.

</blockquote>

## Usage

censusaggregate has two primary functions, the first of which is to retrieve a census vector and *all* of its child vectors (children). Child vectors further break down or provide additional detail to a vector. For example, if we want to get the vector that describes the structural type of dwellings, first we list all vectors for the 2021 census, then filter for the one of interest:

```{r}
library(cancensus)
library(censusaggregate)
library(dplyr)

dataset <- "CA21"

vector <- list_census_vectors(dataset) %>%
  filter(label == "Occupied private dwellings by structural type of dwelling data")

vector
```

Next we can retrieve the vector using its code (`r vector[["vector"]]`) and its children, for a given set of regions and at a given level. Here, we retrieve data for three Census subdivisions (CSD), then look at the data, which includes information like the highest parent vector, a description of the vector, the units its in, its direct parent (since children can be nested), the geographic identifier of the CSD, the population and number of households in that CSD, and the data value itself.

```{r}
vector <- vector[["vector"]]

csd_regions <- list(CSD = c("3520005", "3521005", "3521010"))

vector_with_breakdowns <- get_census_vectors_and_children(
  dataset,
  regions = csd_regions,
  level = "CSD",
  vectors = vector
)

vector_with_breakdowns
```

We can see that there are 10 vectors returned - 9 breakdowns and the original parent vector - each of which has 3 records, one for each of the CSDs we requested data for.

```{r}
vector_with_breakdowns %>%
  count(vector, parent_vector, label)
```

And we can quickly look at the values themselves:

```{r}
vector_with_breakdowns %>%
  select(vector, label, geo_uid, value)
```

The second function of censusaggregate is to aggregate the census vectors across multiple geographies. In this case, we want to aggregate each variable so that there is only one record for each, combining the values across the three CSDs. To do this, we use `aggregate_census_vectors`:

```{r}
vector_with_breakdowns %>%
  aggregate_census_vectors() %>%
  select(vector, label, value, value_proportion)
```

In this case, it returns a field `value` which is the sum of the values within the three CSDs, and `value_proportion`, which breaks down the value of the highest parent vector (`r vector`). For example, there are 457,170 households in total a Single-detached house in the queried CSDs, which represents 28.8% of all households in those CSDs.

## Supported vector aggregation

The vector aggregation happens automatically, based on the vectors' units and aggregation type:

```{r}
vectors <- list_census_vectors(dataset) %>%
  derive_aggregation_type()

vectors  %>%
  count(units, aggregation_type)
```

In the previous example, we worked with vectors with `units = "Number"` and `aggregation_type = "Additive"`, which is one of the three supported aggregations:

1. Units: Number, Aggregation Type: Additive
2. Units: Percentage (0-100), Aggregation Type: Average
3. Units: Currency, Aggregation Type: Average

### Units: Number, Aggregation Type: Additive

The `value` is derived by summing all of the `value` for each vector, across geographies. The result is a count (e.g. population, households). An additional field, `value_proportion`, is derived, describing what proportion of the total count (e.g. of population, households) from the highest parent vector falls within each child vector.

### Units: Percentage (0-100), Aggregation Type: Average

### Units: Currency, Aggregation Type: Average

### Special cases

There are a few vectors that require use of special functions instead of the automatic `aggregate_census_vectors`.

#### aggregate_population_change()

To aggregate Population Change across geographies, `aggregate_population_change` is used on data that contains two population vectors:

```{r}
population_vectors <- vectors %>% 
  filter(label %in% c("Population, 2021", "Population, 2016")) %>%
  pull(vector)

population_change_data <- get_census_vectors_and_children(
  dataset = dataset,
  regions = csd_regions,
  level = "CSD",
  vectors = population_vectors
)

population_change_data %>%
  select(vector, label, geo_uid, value)

population_change_data %>%
  aggregate_population_change()
```

#### aggregate_population_density()

Similarly, to aggregate Population Density across geographies, `aggregate_population_density` is used, passing data that contains a population vector and the Land area in square kilometres vector:

```{r}
population_density_vectors <- vectors %>%
  filter(label %in% c("Population, 2021", "Land area in square kilometres")) %>%
  pull(vector)

population_density_data <- get_census_vectors_and_children(
  dataset = dataset,
  regions = csd_regions,
  level = "CSD",
  vectors = population_density_vectors
)

population_density_data %>%
  select(vector, label, geo_uid, value)

population_density_data %>%
  aggregate_population_density()
```

#### aggregate_estimated_median_income()

We can also aggregate an estimated median income across geographies using `aggregate_estimated_median_income()`. The method used looks at the buckets of income that are available and locates which bucket contains the median value, then assumes a uniform distribution within that bucket to locate the estimated median value. It is then rounded to the nearest $1,000. Note that the median is not calculated if the total count is less than 1,000, nor if the median is in the largest bucket (both cases reutnr NA).

```{r}
total_household_income_vector <- vectors %>%
  filter(label == "Household total income groups in 2020 for private households") %>%
  pull(vector)

total_household_income_data <- get_census_vectors_and_children(
  dataset = dataset,
  regions = csd_regions,
  level = "CSD",
  vectors = total_household_income_vector,
  terminal_only = TRUE # Pulling the data with terminal vectors only, e.g. not $100,000 and over, but its children
)

total_household_income_data %>%
  select(vector, label, geo_uid, value)
```

First, use `aggregate_census_vectors()` on the data to get one record per vector (as opposed to one per geography and vector), then use the estimated median function:

```{r}
total_household_income_data %>%
  aggregate_census_vectors() %>%
  aggregate_estimated_median_income()
```

### Unsupported aggregations

The following are not supported:

1. Units: Number, Aggregation Type: Average
1. Units: Number, Aggregation Type: Median
1. Units: Currency, Aggregation Type: Median
1. Units: Ratio, Aggregation Type: Average
1. Units: Percentage (0-100), Aggregation Type: Median

If you try to aggregate this data, a message and empty tibble will be returned:

```{r}
median_vector <- vectors %>%
  filter(units == "Currency", aggregation_type == "Median") %>%
  slice(1) %>%
  pull(vector)

get_census_vectors_and_children(
  dataset,
  regions = list(CSD = c("3520005", "3521005", "3521010")),
  level = "CSD",
  vectors = median_vector
) %>%
  aggregate_census_vectors()
```

## Utility functions

### collapse_census_vectors()

### derive_aggregation_type()

### derive_census_vector_order()

### reassign_parent_vector()

## Visualizing census data

## plot_census_vector()

## inline_barchart

## Related work

censusaggregate heavily relies on existing work from [`cancensus`]() (and the related [CensusMapper]()) to fetch census data. It has some similar functionality to the [`tongfen`]() package, but is not as feature rich - rather, it focuses on exclusively getting and aggregating Canadian census vectors across multiple (shared) geographies (in contrast to `tongfen`, which ...)
