---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# censusaggregate

The goal of censusaggregate is to easily retrieve and aggregate Canadian census vectors, and their breakdowns, across multiple (shared) geographies.

## Installation

You can install censusaggregate from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("purposeanalytics/census-aggregator-package")
```

## API key and cacheing

TODO

## Usage

```{r}
library(cancensus)
library(censusaggregate)
library(dplyr)

dataset <- "CA16"

vector <- list_census_vectors(dataset) %>%
  filter(label == "Occupied private dwellings by structural type of dwelling data")

vector

vector <- vector[["vector"]]

vector_with_breakdowns <- get_census_vectors_and_children(
  dataset,
  regions = list(CSD = c("3520005", "3521005", "3521010")),
  level = "CSD",
  vectors = vector
)

vector_with_breakdowns %>%
  glimpse()

vector_with_breakdowns %>%
  aggregate_census_vectors() %>%
  glimpse()
```

## Related work

censusaggregate heavily relies on existing work from [`cancensus`]() (and the related [CensusMapper]()) to fetch census data. It has some similar functionality to the [`tongfen`]() package, but is not as feature rich - rather, it focuses on exclusively getting and aggregating Canadian census vectors across multiple (shared) geographies (in contrast to `tongfen`, which ...)
