---
output: html_document
---

```{r, echo = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr, warn.conflicts = FALSE)
library(censusaggregate)
library(purrr)
```

```{r data}
regions <- list(CSD = c("3520005", "3521005", "3521010"))
level <- "CSD"

variables <-
  tibble::tribble(
    ~vector, ~vector_label, ~vector_label_short,
    "v_CA16_401", "Population, 2016", "population_2016",
    "v_CA16_402", "Population, 2011", "population_2011",
    "v_CA16_407", "Land area in square kilometres", "area",
    "v_CA16_1", "Total - Age", "age",
    "v_CA16_418", "Private households by household size", "household_size",
    "v_CA16_484", "Total number of census families in private households - 100% data", "family_type",
    "v_CA16_504", "Total - Private households by household type - 100% data", "household_type",
    "v_CA16_512", "Total - Knowledge of official languages for the total population excluding institutional residents - 100% data", "knowledge_of_english",
    "v_CA16_1355", "Total - Language spoken most often at home for the total population excluding institutional residents - 100% data", "top_10_languages",
    "v_CA16_2525", "In low income based on the Low-income measure, after tax (LIM-AT)", "lim_at",
    "v_CA16_3405", "Total - Immigrant status and period of immigration for the population in private households - 25% sample data", "immigrant_status",
    "v_CA16_3852", "Total - Aboriginal identity for the population in private households - 25% sample data", "aboriginal_identity",
    "v_CA16_3954", "Total - Visible minority for the population in private households - 25% sample data", "visible_minority",
    "v_CA16_3999", "Total - Ethnic origin for the population in private households - 25% sample data", "ethnic_origin",
    "v_CA16_4836", "Total - Private households by tenure - 25% sample data", "household_tenure",
    "v_CA16_4886", "Total -  Owner and tenant households with household total income greater than zero, in non-farm, non-reserve private dwellings by shelter-cost-to-income ratio - 25% sample data", "household_rent?",
    "v_CA16_5051", "Total - Highest certificate, diploma or degree for the population aged 15 years and over in private households - 25% sample data", "educational_attainment"
  )

# Get data for variables and children ----

data <- get_census_variables_and_children(
  regions = regions,
  level = level,
  variables = variables[["vector"]]
)
```

```{r data_export}
# Prep data for export
# Choose which vectors to keep within each

breakdown_vectors <- list(
  "v_CA16_401" = NA,
  "v_CA16_418" = c("v_CA16_419", "v_CA16_420", "v_CA16_421", "v_CA16_422", "v_CA16_423"),
  "v_CA16_3954" = c(
    "v_CA16_3957", "v_CA16_3960", "v_CA16_3963", "v_CA16_3966",
    "v_CA16_3969", "v_CA16_3972", "v_CA16_3975", "v_CA16_3978", "v_CA16_3981",
    "v_CA16_3984", "v_CA16_3987", "v_CA16_3990", "v_CA16_3993",
    "v_CA16_3996"
  ),
  "v_CA16_4836" = c("v_CA16_4837", "v_CA16_4838", "v_CA16_4839"),
  "v_CA16_1" = c(
    "v_CA16_7", "v_CA16_25", "v_CA16_43", "v_CA16_64", "v_CA16_82",
    "v_CA16_100", "v_CA16_118", "v_CA16_136", "v_CA16_154", "v_CA16_172",
    "v_CA16_190", "v_CA16_208", "v_CA16_226", "v_CA16_247", "v_CA16_265",
    "v_CA16_283", "v_CA16_301", "v_CA16_322", "v_CA16_340", "v_CA16_358", "v_CA16_376"
  ),
  "v_CA16_512" = c("v_CA16_515", "v_CA16_521"),
  "v_CA16_3405" = c("v_CA16_3411", "v_CA16_3432"),
  "v_CA16_3852" = c("v_CA16_3855"),
  "v_CA16_5051" = c("v_CA16_5054", "v_CA16_5057", "v_CA16_5060"),
  "v_CA16_484" = c("v_CA16_485", "v_CA16_488", "v_CA16_489"),
  "v_CA16_504" = c("v_CA16_505", "v_CA16_508", "v_CA16_509")
)

filter_breakdown_vectors <- function(data, vector) {
  
  vector_breakdown_vectors <- breakdown_vectors[[vector]]
  
  if (!all(is.na(vector_breakdown_vectors))) {
    data <- data %>%
      filter(
        vector %in% c(vector_breakdown_vectors, !!vector)
      )
  } else {
    data <- data %>%
      filter(
        vector == !!vector
      )
  }

  data
}

# Aggregate vectors - treat v_CA16_407 (land area) separately, since it will be used for population density, and v_CA16_402 (population 2011) separately, since it will be used for population change

variables <- variables %>%
  filter(!(vector %in% c("v_CA16_407", "v_CA16_402")))

data_breakdown <- map_dfr(
  variables[["vector"]],
  function(vector) {
    data %>%
      filter_breakdown_vectors(vector) %>%
      aggregate_census_variables()
  }
) %>%
  select(highest_parent_vector, label, value, value_proportion) %>%
  distinct() %>%
  left_join(variables, by = c("highest_parent_vector" = "vector")) %>%
  select(highest_parent_vector, vector_label_short, label, value, value_proportion)
```


# Population

```{r population}
data_breakdown %>%
  filter(vector_label_short == "population_2016") %>%
  select(label, value, value_proportion)
```

# Number of households

```{r}
data %>%
  select(geo_uid, households) %>%
  distinct() %>%
  summarise(across(households, sum, na.rm = TRUE))
```

# Population change (2011 to 2016)

```{r population_change}
data %>%
  aggregate_population_change()
```

# Population density

```{r population_density}
data %>%
  filter(label %in% c("Population, 2016", "Land area in square kilometres")) %>%
  aggregate_population_density()
```   

# Private households by household size

```{r household_size}
data_breakdown %>%
  filter(vector_label_short == "household_size") %>%
  select(label, value, value_proportion)
```

# Estimated median household income

# Unaffordable housing

# Low-income measure after tax (LIM-AT)

# Visible minority population

```{r visible_minority}
data_breakdown %>%
  filter(vector_label_short == "visible_minority") %>%
  select(label, value, value_proportion)
```

# Household tenure

```{r household_tenure}
data_breakdown %>%
  filter(vector_label_short == "household_tenure") %>%
  select(label, value, value_proportion)
```

# Average shelter cost

# Age 

```{r age}
data_breakdown %>%
  filter(vector_label_short == "age") %>%
  select(label, value, value_proportion)
```

# Language spoken at home

# Knowledge of English

```{r knowledge_of_english}
data_breakdown %>%
  filter(vector_label_short == "knowledge_of_english") %>%
  select(label, value, value_proportion)
```

# Immigrant status

```{r immigrant_status}
data_breakdown %>%
  filter(vector_label_short == "immigrant_status") %>%
  select(label, value, value_proportion)
```

# Aboriginal identity

```{r aboriginal_identity}
data_breakdown %>%
  filter(vector_label_short == "aboriginal_identity") %>%
  select(label, value, value_proportion)
```

# Country of origin

# Educational attainment

```{r educational_attainment}
data_breakdown %>%
  filter(vector_label_short == "educational_attainment") %>%
  select(label, value, value_proportion)
```

# Family type

```{r family_type}
data_breakdown %>%
  filter(vector_label_short == "family_type") %>%
  select(label, value, value_proportion)
```

# Multiple family households

```{r household_type}
data_breakdown %>%
  filter(vector_label_short == "household_type") %>%
  select(label, value, value_proportion)
```
